### 2017-08-17 paper.tex
Update content and add multiple styles.

% ACM
\documentclass[sigconf]{acmart}

\usepackage{booktabs} % For formal tables

\input{paper-preamble.tex}

% Copyright
%\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
%\acmDOI{10.475/123_4}

% ISBN
%\acmISBN{123-4567-24-567/08/06}

%Conference
\acmConference[WOODSTOCK'97]{ACM Woodstock conference}{July 1997}{El Paso, Texas USA} 
\acmYear{1997}
\copyrightyear{2016}

%\acmPrice{15.00}

\begin{document}
\title{Network Structure, Efficiency, and Performance in WikiProjects}

% ACM
\author{Edward L. Platt}
\affiliation{%
  \institution{University of Michigan}
  \city{Ann Arbor} 
  \state{Michigan} 
}
\email{elplatt@umich.edu}
\author{Daniel M. Romero}
\affiliation{%
  \institution{University of Michigan}
  \city{Ann Arbor} 
  \state{Michigan} 
}
\email{drom@umich.edu}

% The default list of authors is too long for headers}
\renewcommand{\shortauthors}{E.L. Platt \& D.M. Romero}

% ACM
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%
\begin{CCSXML}
\end{CCSXML}

% ACM
\ccsdesc[500]{Computer systems organization~Embedded systems}
\ccsdesc[300]{Computer systems organization~Redundancy}
\ccsdesc{Computer systems organization~Robotics}
\ccsdesc[100]{Networks~Network reliability}

% ACM
\keywords{collaboration, cooperation, peer production, wikipedia, wikiprojects, }

\input{paper-content.tex}

% ACM
\bibliographystyle{ACM-Reference-Format}

\bibliography{paper}

\end{document}